---
title: 02_DE
author: Qian Hui TAN
date: '`r format(Sys.time(), "%d %B %Y")`'
format:
  html: 
    keep-md: true
    embed-resources: true
    df-print: kable
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
    smooth-scroll: true
    code-tools: true
    code-line-numbers: true
  gfm: 
    df-print: kable
    toc: true 
    toc-depth: 3
    number-sections: true
    prefer-html: true
---

# 02_DE

In this notebook, we run quick DE volcano plots, p-value histograms and pairwise cross plots for our conditions. 



```{r}
#| label = "setup",
#| include = FALSE

knitr::opts_chunk$set(fig.path='../figures/02_DE/',
                      dev=c("png"))

```


```{r}
source("../../scripts/R/de_functions.R")
source("../../scripts/R/go_kegg_functions.R")

suppressPackageStartupMessages({
  library(DESeq2)
  library(GenomicFeatures)
  library(RColorBrewer)
  library(gplots)
  library(biomaRt)
  library(grid)
  library(gridExtra)
  library(ggplot2)
  library(lattice)
  library(reshape)
  library(geneplotter)
  library(ggrepel)
  library(limma)
  library(tidyverse)
  library(eulerr)
  library(gghighlight)
  library(clusterProfiler) # install from github
  library(org.Dm.eg.db)
})


```


## Functions


```{r}
## If the output directory doesn't exist, create it
if(!dir.exists("../output/02_DE")){
  dir.create("../output/02_DE")
}

output_dir <- "../output/02_DE/"
```

Loading data: 

```{r}
#| label = "load data"

dds = readRDS("../output/01_QC/dds_filt.RDS")
ensembl.genes= readRDS("../output/01_QC/ensembl_genes.RDS")
lfc.threshold = log2(1.5)


dim(dds)
```


Remove lowly expressed genes: 
```{r}
# Set control as reference
#colData(dds)$condition <- relevel(colData(dds)$condition, ref = "WT")


# Removing lowly expressed genes, only to be done once at the start of the differential expression step
filter = apply(counts(dds, normalized = TRUE), 1, function(x){ mean(x) >= 10 })
dds = dds[filter, ]
dim(dds)
```

Run DESeq: 
```{r}
dds <- DESeq(dds, test = "Wald", parallel = TRUE)
rld <- rlog(dds, blind = FALSE) # blind = FALSE for downstream DE

design(dds) 
```

List of samples in here
```{r}
colData(dds)$sample_id
```



## RLE plot
```{r}
#| label: rle-plot

assay(rld) %>%
  as.data.frame() %>%
  gather() %>%
  ggplot(aes(x=key, y=value)) + 
  geom_boxplot() + ylab("Log Expression ") + xlab("Sample") +
  theme(axis.text.x = element_text(angle = 90))
```

Run Wald test
```{r}
wald_dds = nbinomWaldTest(dds)
```

Check names of coefficients
```{r}
res_names <- resultsNames(wald_dds)

res_names
```


# Wald Tests - all

In this section, we export all the pairwise comparisons (W4_OE vs Ctrl, O1_OE vs Ctrl, etc) as an excel sheet. Each sheet will contain one comparison. 
```{r}
# Function to extract results from a given comparison 
get_res <- function(x) {
  res = results(wald_dds, name = x) 
  res$ensembl_gene_id = rownames(res)
  res$external_gene_name = ensembl.genes$external_gene_name[match(rownames(res), ensembl.genes$ensembl_gene_id)]
  res$gene_biotype = ensembl.genes$gene_biotype[match(rownames(res), ensembl.genes$ensembl_gene_id)]
  
  res <- as_tibble(res) %>% 
    dplyr::select(ensembl_gene_id, log2FoldChange, pvalue, padj, gene_biotype, external_gene_name)
  
  return(res)
  }

wald_res_all <- lapply(res_names, get_res)
names(wald_res_all) <- res_names

head(wald_res_all)
```


```{r}
# export this
openxlsx::write.xlsx(wald_res_all, 
                     file = paste0(output_dir, "logfoldchanges.xlsx"), 
                     sheetName = names(wald_res_all), 
                     rowNames = FALSE)
```


# Wald Tests

## W4_OE
:::{.panel-tabset}

### Histogram

```{r w4oe_ctrl}
#| label: W4OE_CTRL
### Note the order of W4_OE and CTRL - W4_OE here is the numerator, CTRL is the denominator

results_W4OE_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "W4_OE", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_W4OE_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: W4OE-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: W4OE-WT-vlc
### Volcano plot
generate_volcano(res, "W4_OE", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")
```

```{r}
ego
```

### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")
```

```{r}
ego
```


### KEGG, up

```{r}
#| label: up-ekegg

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
write.csv(up_ekegg, file = paste0(output_dir, "w4oe_vs_ctrl_up_kegg.csv"))
```

### KEGG - down

```{r}
#| label: down-ekegg
#| fig-height: 10

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
write.csv(down_ekegg, file = paste0(output_dir, "w4oe_vs_ctrl_down_kegg.csv"))
```

:::


## O1_OE
:::{.panel-tabset}

### Histogram

```{r o1oe_ctrl}
#| label: O1OE_CTRL
### Note the order of W4_OE and CTRL - W4_OE here is the numerator, CTRL is the denominator

results_O1OE_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O1_OE", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O1OE_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```



### MA plot

```{r}
#| label: O1OE-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O1OE-CTRL-vlc
### Volcano plot
generate_volcano(res, "O1_OE", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: o1oe-up-go
ego_up <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")
```

```{r}
#ego_up
```


### Custom GO, up
```{r}
#| label: o1oe-custom-go-up

interesting_pathways <- c("nuclear division", 
                          "regulation of cell cycle",
                          "blastoderm segmentation",
                          "embryonic pattern specification",
                          "anterior/posterior axis specification", 
                          "nuclear chromosome segregation", 
                          "mitotic cell cycle",
                          "DNA repair", 
                          "mismatch repair",
                          "spindle organization")

# Takes in an ego tibble, subsets interesting pathways. 
custom_ego_table(ego_up, interesting_pathways)


custom_ego(ego_up, interesting_pathways,
           title = "GO, O1_OE vs Ctrl")
```

### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: o1oe-go-down
ego_down <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")
```
### Custom GO, down
```{r}
#| label: o1oe-custom-go-down
interesting_pathways <- c("DNA replication", 
                          "DNA repair",
                          "recombinational repair",
                          "double-strand break repair")

custom_ego(ego_down, interesting_pathways, 
           title = "GO, downregulated, O1OE vs Ctrl")

custom_ego_table(ego_down, interesting_pathways)
```

### KEGG, up

```{r}
#| label: o1oe-up-ekegg

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

### Custom KEGG, up
```{r}
#| label: o1oe-kegg-up
custom_ekegg(up_ekegg, interesting_pathways = up_ekegg$Description,
             title = "KEGG, up, O1OE vs Ctrl")
```


```{r}
write.csv(up_ekegg, file = paste0(output_dir, "o1oe_vs_ctrl_up_kegg.csv"))
```

### KEGG - down

```{r}
#| label: o1oe-kegg-down

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```


```{r}
down_ekegg
```
### Custom KEGG, down

```{r}
#| label: o1oe-kegg-custom-down

interesting_pathways = c("DNA replication",
                         "Base excision repair",
                         "Mismatch repair",
                         "Nucleotide excision repair")

custom_ekegg(down_ekegg, 
             interesting_pathways, 
             title = "KEGG, down, O1OE vs Ctrl")
```


```{r}
write.csv(down_ekegg, file = paste0(output_dir, "o1oe_vs_ctrl_down_kegg.csv"))
```

:::


## O1W4_OE
:::{.panel-tabset}

### Histogram

```{r o1w4oe_ctrl}
#| label: O1W4OE_CTRL

results_O1W4OE_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O1W4_OE", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O1W4OE_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O1W4OE-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O1W4OE-WT-vlc
### Volcano plot
generate_volcano(res, "O1W4_OE", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O1W4OE-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")
```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "o1w4oe_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O1W4OE-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")
```

```{r}
nrow(ego)

write.csv(ego, file = paste0(output_dir, "o1w4oe_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: O1W4OE-up-ekegg

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "o1w4oe_vs_ctrl_up_kegg.csv"))
```


```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: O1W4OE-down-ekegg
#| fig-height: 10

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "o1w4oe_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::


## O2_KO
:::{.panel-tabset}

### Histogram

```{r O2KO_ctrl}
#| label: O2KO_CTRL

results_O2KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O2_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O2KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O2KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O2KO-WT-vlc
### Volcano plot
generate_volcano(res, "O2_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O2KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O2KO_vs_ctrl_down_go.csv"))
```

### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O2KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O2KO_vs_ctrl_down_go.csv"))
```


### KEGG, up

```{r}
#| label: O2KO-up-ekegg

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "O2KO_vs_ctrl_up_kegg.csv"))
```

### KEGG - down

```{r}
#| label: O2KO-down-ekegg

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "O2KO_vs_ctrl_down_kegg.csv"))
```

:::


## W4_KO
:::{.panel-tabset}

### Histogram

```{r W4KO_ctrl}
#| label: W4KO_CTRL

results_W4KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "W4_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_W4KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: W4KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: W4KO-WT-vlc
### Volcano plot
generate_volcano(res, "W4_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: W4KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "W4KO_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: W4KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "W4KO_vs_ctrl_down_go.csv"))
```


```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: W4KO-up-ekegg
#| fig-height: 10

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "W4KO_vs_ctrl_up_kegg.csv"))
```

```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: W4KO-down-ekegg

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "W4KO_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::


## O1_KO
:::{.panel-tabset}

### Histogram

```{r O1KO_ctrl}
#| label: O1KO_CTRL

results_O1KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O1_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O1KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O1KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O1KO-WT-vlc
### Volcano plot
generate_volcano(res, "O1_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O1KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1KO_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O1KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1KO_vs_ctrl_down_go.csv"))
```


```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: O1KO-up-ekegg
#| fig-height: 10

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "O1KO_vs_ctrl_up_kegg.csv"))
```

```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: O1KO-down-ekegg
#| fig-height: 10

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "O1KO_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::


## O1W4_KO
:::{.panel-tabset}

### Histogram

```{r O1W4KO_ctrl}
#| label: O1W4KO_CTRL

results_O1W4KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O1W4_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O1W4KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O1W4KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O1W4KO-WT-vlc
### Volcano plot
generate_volcano(res, "O1W4_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O1W4KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1W4KO_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O1W4KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1W4KO_vs_ctrl_down_go.csv"))
```


```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: O1W4KO-up-ekegg
#| fig-height: 10

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "O1W4KO_vs_ctrl_up_kegg.csv"))
```

```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: O1W4KO-down-ekegg

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "O1O1W4KO_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::


## O2W4_KO
:::{.panel-tabset}

### Histogram

```{r O2W4KO_ctrl}
#| label: O2W4KO_CTRL

results_O2W4KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O2W4_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O2W4KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O2W4KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O2W4KO-WT-vlc
### Volcano plot
generate_volcano(res, "O2W4_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O2W4KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O2W4KO_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O2W4KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O2W4KO_vs_ctrl_down_go.csv"))
```


```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: O2W4KO-up-ekegg
#| fig-height: 10

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "O2W4KO_vs_ctrl_up_kegg.csv"))
```

```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: O2W4KO-down-ekegg

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "O2W4KO_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::


## O1O2_KO
:::{.panel-tabset}

### Histogram

```{r O1O2KO_ctrl}
#| label: O1O2KO_CTRL

results_O1O2KO_CTRL = get_dds_res(wald_dds, 
                             contrast = c("condition", "O1O2_KO", "Ctrl"),
                             ensembl.genes, 
                             shrink = TRUE)

res = results_O1O2KO_CTRL 

hist(res$pvalue, xlab = "p-value", ylab = "Frequency")
```

**Number of significant genes (padj < 0.1 & log2FoldChange < log2(1.5)):**

```{r}
sum(res$padj < 0.1 & abs(res$log2FoldChange) > lfc.threshold, na.rm = T)
```


### MA plot

```{r}
#| label: O1O2KO-CTRL-MA

### MA plot
generate_ma(res)
```

### Volcano plot

```{r}
#| label: O1O2KO-WT-vlc
### Volcano plot
generate_volcano(res, "O1O2_KO", "Ctrl")
```
### GO - up

```{r}
df_res = as.data.frame(res)
# Drop rows with NAs in pvalue - these are genes with high variability
df_res = df_res[complete.cases(df_res[ ,c("pvalue", "padj")]), ] 
```


```{r}
up_genes <- rownames(res[res$padj < 0.1 &
                      res$log2FoldChange > lfc.threshold, ])

length(up_genes)
```

```{r}
#| label: O1O2KO-up-go
ego <- plotEGO_dm(up_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, upregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1O2KO_vs_ctrl_down_go.csv"))
```

```{r}
head(ego)
```


### GO - down

```{r}
down_genes <- rownames(df_res[df_res$padj < 0.1 &
                      df_res$log2FoldChange < lfc.threshold, ])

length(down_genes)
```


```{r}
#| label: O1O2KO-down-go
ego <- plotEGO_dm(down_genes, universe = rownames(dds), 
               ont = "BP",
               title = "GO, downregulated genes")

```

```{r}
nrow(ego)
write.csv(ego, file = paste0(output_dir, "O1O2KO_vs_ctrl_down_go.csv"))
```


```{r}
head(ego)
```


### KEGG, up

```{r}
#| label: O1O2KO-up-ekegg
#| fig-height: 10

# Get the entrez IDs
up_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% up_genes, ]$entrezgene_id)

up_ekegg <- plotKEGG_dm(up_entrez, title = "KEGG, upregulated genes")
```

```{r}
nrow(up_ekegg)
write.csv(up_ekegg, file = paste0(output_dir, "O1O2KO_vs_ctrl_up_kegg.csv"))
```

```{r}
head(up_ekegg)
```


### KEGG - down

```{r}
#| label: O1O2KO-down-ekegg

# Get the entrez IDs
down_entrez <- na.omit(ensembl.genes[ensembl.genes$gene_id %in% down_genes, ]$entrezgene_id)

down_ekegg <- plotKEGG_dm(down_entrez, title = "KEGG, downregulated genes")
```

```{r}
nrow(down_ekegg)
write.csv(down_ekegg, file = paste0(output_dir, "O1O2KO_vs_ctrl_down_kegg.csv"))
```

```{r}
head(down_ekegg)
```


:::




# sessionInfo



```{r}
sessionInfo()
```

